---
title: "BiocDockerManager Vignette"
author: "Nitesh Turaga"
package: BiocDockerManager
output: 
    BiocStyle::html_document
vignette: >
    %\VignetteIndexEntry{BiocDockerManager Vignette}
    %\VignetteEngine{knitr::rmarkdown}
    %\VignetteEncoding{UTF-8}  
---

## Introduction

The `BiocDockerManager` package works analogous to BiocManager but for
docker images. Use the BiocDockerManager package to install and manage
docker images provided by the Bioconductor project. The package
provides convienient ways to install images, update images, and find
which Bioconductor based docker images are available.

## Preliminaries 

### Installation

If you are reading this document and have not yet installed any
software on your computer, visit http://bioconductor.org and follow
the instructions for installing R and Bioconductor. Once you have
installed R and Bioconductor, you are ready to go with this document.
In the future, you might find that you need to install one or more
additional packages. The best way to do this is to start an R session
and evaluate commands like

```{r}
   if (!require("BiocManager"))
      install.packages("BiocManager")
   BiocManager::install("BiocDockerManager")
```

We require the system to have the docker engine installed. It is
supported on all platforms. Please take a look at
[https://docs.docker.com/install/](https://docs.docker.com/install/).

### Loading packages

```{r}
library("BiocDockerManager")
```

## Usage

The best way to describe the package and it's contents are to showcase
some of the functions available in this package. A few functions use
the the REST API provided by https://hub.docker.com/, and a few of
them interact with the `docker` command line tool installed on your
local machine.

### Availability of images

First we can check what Bioconductor docker images are
"available". This returns a tibble which we can manipulate with basic
`dplyr` functionality.

```{r available, echo=TRUE}
library(dplyr)
res <- BiocDockerManager::available()

res

res2 <- BiocDockerManager::available(pattern = "bioconductor_docker")

res2

res2 %>% select(IMAGE, TAGS)
```

Bioconductor also has a list of images which have been recently
deprecated,

```{r deprecated, echo=TRUE}
BiocDockerManager::available(deprecated=TRUE)
```

### Help regarding Bioconductor images

The custom help function takes users to the Bioconductor Docker help
page. This provides easy access to the help page
should they need it. The help page contains all the information
needed on how to start up the `bioconductor/bioconductor_docker`
image, and use it.

```{r, eval=FALSE}
BiocDockerManager::help()
```

### Install or "pull" a docker image

The `install()` function lets users download and pull new images 
on to their machine. This function comes in handy when users are 
trying to get a new `bioconductor_docker` image on their machine.

The function pulls the image from **Dockerhub** which is always
kept upto date.

```{r, eval = FALSE}
BiocDockerManager::install("bioconductor/bioconductor_docker", 
                            tag = "latest")
```

### Installed images on local machine

This function allows the users to check which images are installed
on their local machine. This is similar to checking with the docker
command line function `docker images`.

For this function to work you need to have the Docker engine running.

```{r}
BiocDockerManager::installed()
```

You can also filter the list of installed images by providing the 
`name` arguement

```{r}
BiocDockerManager::installed(name = "bioconductor/bioconductor_docker")
```

### Image details 

Bioconductor developed and distributed images require that each
Dockerfile which accompanies the image has LABEL(s) which identify
the maintainer, version, description, url and a few other entities. 

The following functions allow to query the Docker image for these 
LABEL identities.

These labels serve a similar purpose to the DESCRIPTION file in a 
Bioconductor package.

Since these functions are specific to Bioconductor Docker images,
they will not work when given generic docker images.

```{r}
## Get version for the "latest" tag bioconductor image
BiocDockerManager::maintainer(name = "bioconductor/bioconductor_docker", tag = "latest")

## The above functions works the same as
BiocDockerManager::maintainer(tag = "latest")
```

The version number of the Docker image is very informative as it tells 
us the Bioconductor version number in the `x.y` coordinates, and the `.z`
informs us the version of the Dockerfile which is used to build the
Docker image.

```{r}
BiocDockerManager::version(name = "bioconductor/bioconductor_docker", tag = "latest")

## Get image version
BiocDockerManager::version(tag = "latest")

BiocDockerManager::version(tag = "devel")

BiocDockerManager::version(tag = "RELEASE_3_10")
```

### Check validity of images

It is always a good idea to check if the images you are running are 
the latest ones produced by Bioconductor and if they are up to date.

The `valid()` function takes the **digest** of the docker image you 
have locally and the docker image hosted on Dockerhub to check if the 
digest SHA's match. 

If the docker image is not updated, you might want to `install()` the 
latest version of the Docker image. 

The function returns a tibble with the images that need to be updated.
The tibble contains image name and tag.

```{r}
BiocDockerManager::valid(name = "bioconductor/bioconductor_docker", tag = "latest")
```

### Dockerfile template for Bioconductor images

This `use_dockerfile()` functionality is for people who want to 
contribute Docker images to the Bioconductor ecosystem. The function
has side-effects where it creates a directory and populates it with a
**Dockerfile** and **README.md**. 

The Dockerfile has fields which are required for the submission process
of the Docker image. 

There are best practices and standards on how to contribute new docker
images to Bioconductor. The following link has more information.
http://bioconductor.org/help/docker/#contribute

```{r, eval = FALSE}
BiocDockerManager::use_dockerfile()
```

## Workflow

We hope to provide functionality which is useful to R and Bioconductor
Docker users in the form of an R package.

The typical workflow would look like the following:

```{r}

BiocDockerManager::available()

BiocDockerManager::install()

BiocDockerManager::version("bioconductor/bioconductor_docker", tag = "devel")

BiocDockerManager::valid("bioconductor/bioconductor_docker", tag = "devel")
```

Once some time has passed and if you are not sure if you image is upto date,

```{r}
BiocDockerManager::valid("bioconductor/bioconductor_docker", tag = "devel")

BiocDockerManager::install("bioconductor/bioconductor_docker", tag = "devel")
```


## sessionInfo

```{r}
sessionInfo()
```


